Technical Report

### 2018-06-14 fully-abstract-comp/technical-report.tex
Sketch of a lemma that could make the call case for the FTLR proof work.
First stab at proving Lau's new lemma.

\[
  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
    \begin{array}{l}
      \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \wedge \\
      \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
    \end{array}
            \right\}
\]
We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
% DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
% Where $S\text{ is a separation of $W$}$ means that no two partition share the same ``spatial'' region (whatever that is).
%\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
% Lau 19-01-2018: Added rdata as default.
\lau{RESOLVED check what conditions need to have a ``square'' i.e. need approximation}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))

@@

\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
}
\dominique{TODO: better name than ``reasonable'' for all of this.}
Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if the following implications hold.
\begin{itemize}
\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ then $[\baddr,\eaddr] \# \ta$
\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
\end{itemize}

@@

>>>>

\[
  \lrrg{\trust}(R)(W) = \left\{ \npair{\stpair{\reg}{\reg}} \middle|
    \begin{array}{l}
      \exists S : (\RegName \setminus (\{\pcreg \} \cup R))\fun \World \ldotp \\
      \quad W = \bigoplus_{r \in (\RegName\setminus (\{\pcreg,\rdata \} \cup R))} S(r) \wedge \\
      \quad \forall r \in \RegName \setminus (\{\pcreg \} \cup R)\ldotp\\
      \qquad\npair{\stpair[.]{\src{\reg_S(r)}}{\reg_T(r)}} \in \lrvg{\trust}(S(r))
    \end{array}
            \right\}
\]
We write $\lrr(W)$ to mean $\lrr(\emptyset)(W)$. That is, if we do not need to exclude extra registers, then we simply omit that argument.
% DOMI 18-1-2018: the below remark is obsolete, so I commented it out.
% Where $S\text{ is a separation of $W$}$ means that no two partitionjjjjshare the same ``spatial'' region (whatever that is).
%\dominique{18-1-2018: it might be convenient to exclude $\rdata$ from the $\lrr$ relation by default too. Check if we ever use it with other instantiations of R.}
% Lau 19-01-2018: Added rdata as default.
\lau{RESOLVED check what conditions need to have a ``square'' i.e. need approximation}
\[
  \lrv(W) =
  \begin{array}[t]{l}
    \left\{ \npair{\stpair[.]{i}{i}} \;\middle|\; i \in \ints \right\}\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l}\src{\sealed{\sigma,\vsc_S}},\\ \sealed{\sigma,\vsc_T} \endarray\right)} \;\middle| \;
    \begin{array}{l}
      (\isLinear{\src{\vsc_S}} \text{ iff } \isLinear{\vsc_T}) \wedge\\
      \exists r \in \dom(\pwheap)\ldotp \pwheap(r) = (\pure,\_,H_\sigma) \tand \\
      \quad \npair[n']{\stpair[.]{\vsc_S}{\vsc_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W) \text{ for all $n' < n$}\wedge\\
      \quad (\isLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad\forall W' \future W, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o)) \wedge \\
      \quad (\nonLinear{\src{\vsc_S}} \Rightarrow \\
      \qquad \forall W' \future \purePart{W}, W_o, n' < n, \npair[n']{\stpair[.]{\vsc_S'}{\vsc'_T}} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \\
      \qquad \quad \npair[n']{\src{\vsc_S},\src{\vsc_S'},\vsc_T,\vsc_T'} \in \lrexj(W'\oplus W_o))

@@

\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
}
\dominique{TODO: better name than ``reasonable'' for all of this.}
Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if the following implications hold.
\begin{itemize}
\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$, then $[\baddr,\eaddr] \# \dom(\mscode)$
\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
\end{itemize}

@@

\begin{lemma}
  \label{lem:codereg-Hsigma-not-lrv}
  If $\dom(\code) \subseteq \ta$, $n > 0$ and $\sigma \in (\sigrets \uplus \sigcloss)$, then $\codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}.H_\sigma~ \sigma \not\nequal \lrv$.
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}
  \label{lem:codereg-not-read-write}
  If $\pwheap(r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$ and $a \in \dom(\code)$, then
  \begin{itemize}
  \item $\npair{\{a\}} \not\in \readCond{\normal,W}$
  \item $\npair{\{a\}} \not\in \writeCond{\normal,W}$
  \end{itemize}
\end{lemma}
\begin{proof}
  TODO
\end{proof}

\begin{lemma}
  If
  \begin{itemize}
  \item $\npair{w,\_} \in \lrv(W_w)$
  \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
  \item $W_w \oplus W_M$ is defined
  \item $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
  \item $\dom(\code) \subseteq \ta$
  \item $n > 0$
  \end{itemize}
  then, with respect to $\code,\ta,\sigrets,\sigcloss$,
  $w$ is reasonable up to $n$ steps in memory $\Phi.\mem$.
\end{lemma}
\begin{proof}
  We need to prove four implications:
  \begin{itemize}
  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$:

    Take $\sigma \in [\sigma_b,\sigma_e]$, we need to prove that $\sigma \not\in (\sigrets \cup \sigcloss)$.
    Assume that $\sigma \in (\sigrets \cup \sigcloss)$ and we will prove a contradiction.

    From $\npair{w,\_} \in \lrv(W_w)$, we get a $r \in \dom(\pwheap[W_w])$ such that $\pwheap[W_w](r) = (\pure,\_,H_\sigma)$ and $H_\sigma~\sigma \nequal \lrv$.

    From $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$, it follows that $\npair{(\overline{\sigma},\_,\_)} \in \lrheap(\pwheap[W_M])(W_{\var{heap}})$ for some $\overline{\sigma}$, $W_{\var{heap}}$ with $W_M = W_{\var{stack}} \oplus W_{\var{free\_stack}} \oplus W_{\var{heap}}$.

    From this, it follows that there exists a $R_\var{seal} : \dom(\activeReg{\pwheap[W_{\var{heap}}]}) \fun \powerset{\Seal}$, $\biguplus_{r \in \dom(\activeReg{\pwheap[W_{\var{heap}}]})} R_\var{seal}(r)) \subseteq \overline{\sigma}$ and $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.

    From $\dom(\pwheap[W_{\var{heap}}](r).H_\sigma) = R_\var{seal}(r)$ and the definedness of all the $\oplus$ operations on worlds, it follows that also $\dom(\pwheap[W_w](r).H_\sigma) = R_\var{seal}(r)$ for all $r$.

    From the facts that $\pwheap[W_w](r_\code) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$, $\pwheap[W_w].H_\sigma~\sigma \nequal \lrv$, the disjointness of all the $R_{\var{seal}}(r)$, it follows that $r = r_\code$.
    However, this contradicts Lemma~\ref{lem:codereg-Hsigma-not-lrv}.

  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$, then $[\baddr,\eaddr] \# \ta$:

    Assume that $a \in [\baddr,\eaddr] \cap \ta$ and we will prove a contradiction.

    From $\npair{w,\_} \in \lrv(W_w)$, it follows that $\npair{\{a\}}\in\readCond{\normal,W_w}$ or $\npair{\{a\}}\in\writeCond{\normal,W_w}$.

    This contradicts Lemma~\ref{lem:codereg-not-read-write}.

  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable:
    TODO

  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$:
    TODO

  \end{itemize}
\end{proof}

\begin{lemma}
  If
  \begin{itemize}
  \item $\npair{\reg_S,\_} \in \lrr(W_w)$
  \item $\memSat{\ms_S,\stk,\ms_\stk,\_}{W_M}$
  \item $W_w \oplus W_M$ is defined
  \item $\pwheap[W_w](r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)}$
  \item $\dom(\code) \subseteq \ta$
  \item $n > 0$
  \end{itemize}
  then, with respect to $\code,\ta,\sigrets,\sigcloss$,
  \begin{itemize}
  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$
  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove the three results separately:
  \begin{itemize}
  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$:

    We have that $W = \bigoplus_{r \in \RegName \setminus \{\pcreg\}} S(r)$ and $\npair{(\reg_S(r),\reg_T(r))} \in \lrv(S(r))$ for all $r \not\in \{\pcreg\}$.

    TODO

  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
    TODO
  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
    TODO
  \end{itemize}
\end{proof}
